### 2016-05-26 syntax.tex
hacking on syntax

\heading{Hash-based (Bloom) filters.}
Fix a nonempty sets $U,\Sigma$ and integers $k,m,n>0$.  Fix a symbol $\bot \not\in U$.  An $(n,k,m)$-filter (over universe~$U$) is a triple  $B=(\mathsc{H},\mathsc{Rep},\mathsc{Qry})$.   
%
The randomized \emph{hash-sampling} algorithm~$\mathsc{H}$ samples a size~$k$ subset~$\mathcal{H} \subset \mathrm{Func}(U,\{0,1,\ldots,m-1\})$.  We write $\mathcal{H} \getsr \mathsc{H}$ for this operation.
%
The deterministic representation algorithm $\mathsc{Rep}\colon [U]^n \rightarrow \Sigma^m$ takes a set~$S$ of~$n$ elements as input, and outputs representation~$M$ of length~$m$.  (Recall that $[U]^n$ denotes subsets of~$U$ of size~$n$.)
%
The determinisitc query algorithm $\mathsc{Qry}\colon \Sigma^m \times U \rightarrow \bits$ takes a representation $M$ and an element $x \in U$ and returns a bit.  
%
We assume that both $\mathsc{Rep}$ and $\mathsc{Qry}$ have the following blackbox access to the functions $h_1,h_2,\ldots,h_k \in \mathcal{H}$, which we denote by writing~$\mathcal{H}$ as a superscript.   Thus, we write $M \gets \mathsc{Rep}^{\mathcal{H}}(S)$ and $b \gets \mathsc{Qry}^{\mathcal{H}}(M,x)$ for the execution of the representation and query algorithms, respectively.
%
For correctness, we demand that for all~$S$ of size~$n$, we have that for all $x \in S, \Prob{\mathsc{Qry}^{\mathcal{H}}(\mathsc{Rep}^{\mathcal{H}}(S),x)=1}$, where the probability is over the sampling $\mathcal{H}\getsr\mathsc{H}$.

@@

Fix a nonempty sets $U,\Sigma$ and integers $k,m_1,m_2,n>0$ with $m_1 \leq m_2$.  Fix a symbol $\bot \not\in U$.  An $(n,k,[m_1,m_2])$-filter (over universe~$U$) is a tuple  $B=(\mathsc{H},\mathsc{Rep},\mathsc{Qry},\mathsc{Up}, \mathsc{Test})$.   
%
The randomized \emph{hash-sampling} algorithm~$\mathsc{H}$ samples a size~$k$ subset~$\mathcal{H} \subset \mathrm{Func}(U,\{0,1,\ldots,m_2-1\})$.  We write $\mathcal{H} \getsr \mathsc{H}$ for this operation.
%
The deterministic representation algorithm $\mathsc{Rep}\colon [U]^n \rightarrow \left(\bigcup_{m=m_1}^{m_2}\Sigma^m\right)$ takes a set~$S$ of~$n$ elements as input, and outputs representation~$M$ of length~$m_1 \leq m \leq m_2$.
%
The determinisitc query algorithm $\mathsc{Qry}\colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right)\times U \rightarrow \bits^*$ takes a representation $M$ and an element $x \in U$ and returns a bitstring.  
%
The deterministic \emph{update} algorithm $\mathsc{Up}\colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right)\times \bits^* \rightarrow \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right) \cup \{\bot\}$ takes a representation and an update string, and returns an updated representation or the distinguished symbol~$\bot$. 
%
The deterministic \emph{test} algorithm $\mathsc{Test} \colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right) \times U \rightarrow \bits$ takes a representation~$M$ and an element~$x \in U$ and returns a bit.
%
We assume that all $\mathsc{Rep},\mathsc{Qry},\mathsc{Up},\mathsc{Test}$ all have blackbox access to the functions $h_1,h_2,\ldots,h_k \in \mathcal{H}$, which we denote by writing~$\mathcal{H}$ as a superscript.

\tsnote{Not sure how to define correctness.  Intuitively, it should be something like, at any point in time, if $x\in S$ then $\Prob{\mathsc{Test}(M,x)=1} \geq \epsilon$ for some correctness parameter~$\epsilon$, where the probablilty is over...?  Coins of~$\mathsc{H}$, perhaps.  (Alternatively, could require for all outputs of~$\mathsc{H}$, as above.)  But perhaps this should be an adversarial game, in which case it is over the coins of the adversary, too.}

@@

An $(n,k,[m_1,m_2])$-filter with operations is a tuple  $B=(\mathsc{H},\mathsc{Rep},\mathsc{Qry}, \mathcal{F})$.  
The set $\mathcal{F}$ is the finite collection of allowable operations.  All operations are of the form 
$f: \mathbb{M}_{\mathcal{U}} \times \mathbb{M}_{\mathcal{U}} \rightarrow \mathbb{M}_{\mathcal{U}} \cup \{\bot\}$.  
%
The deterministic representation algorithm $\mathsc{Rep}\colon \mathbb{M}_\mathcal{U} \rightarrow \bigcup_{m=m_1}^{m_2}\Sigma^m$ takes a multiset~$S$, and outputs representation~$M$ of length~$m_1 \leq m \leq m_2$, or the distinguished symbol~$\bot$.  We assume that if the multiset~$S=\{(x_1,\ell_1),(x_2,\ell_2),\ldots,(x_t,\ell_t)\}$ is such that $n < \sum_{i=1}^t \ell_i$ then $\mathsc{Rep}(S)=\bot$.
%
The randomized hash-sampling algorithm~$\mathsc{H}$ is as before.
%
The determinisitic query algorithm $\mathsc{Qry}$... \tsnote{Not sure how to define this!  See my comment, below...}

>>>>

\heading{Hash-based (Bloom) filters.}
Fix a nonempty sets $U,\Sigma$ and integers $k,m,n>0$.  Fix a symbol $\bot \not\in U$.  An $(n,k,m)$-filter (over universe~$U$) is a triple  $B=(\mathsc{Hash},\mathsc{Rep},\mathsc{Qry})$.   
%
The randomized \emph{hash-sampling} algorithm~$\mathsc{Hash}$ samples a size~$k$ subset~$\mathcal{H} \subset \mathrm{Func}(U,\{0,1,\ldots,m-1\})$.  We write $\mathcal{H} \getsr \mathsc{Hash}$ for this operation.
%
The deterministic representation algorithm $\mathsc{Rep}\colon [U]^n \rightarrow \Sigma^m$ takes a set~$S$ of~$n$ elements as input, and outputs representation~$M$ of length~$m$.  (Recall that $[U]^n$ denotes subsets of~$U$ of size~$n$.)
%
The determinisitc query algorithm $\mathsc{Qry}\colon \Sigma^m \times U \rightarrow \bits$ takes a representation $M$ and an element $x \in U$ and returns a bit.  
%
We assume that both $\mathsc{Rep}$ and $\mathsc{Qry}$ have the following blackbox access to the functions $h_1,h_2,\ldots,h_k \in \mathcal{H}$, which we denote by writing~$\mathcal{H}$ as a superscript.   Thus, we write $M \gets \mathsc{Rep}^{\mathcal{H}}(S)$ and $b \gets \mathsc{Qry}^{\mathcal{H}}(M,x)$ for the execution of the representation and query algorithms, respectively.
%
For correctness, we demand that for all~$S$ of size~$n$, we have that for all $x \in S, \Prob{\mathsc{Qry}^{\mathcal{H}}(\mathsc{Rep}^{\mathcal{H}}(S),x)=1}=1$, where the probability is over the sampling $\mathcal{H}\getsr\mathsc{Hash}$.

@@

Fix a nonempty sets $U,\Sigma$ and integers $k,m_1,m_2,n>0$ with $m_1 \leq m_2$.  Fix a symbol $\bot \not\in U$.  An $(n,k,[m_1,m_2])$-filter (over universe~$U$) is a tuple  $B=(\mathsc{Hash},\mathsc{Rep},\mathsc{Qry},\mathsc{Up}, \mathsc{Test})$.   
%
The randomized \emph{hash-sampling} algorithm~$\mathsc{Hash}$ samples a size~$k$ subset~$\mathcal{H} \subset \mathrm{Func}(U,\{0,1,\ldots,m_2-1\})$.  We write $\mathcal{H} \getsr \mathsc{Hash}$ for this operation.
%
The deterministic representation algorithm $\mathsc{Rep}\colon [U]^n \rightarrow \left(\bigcup_{m=m_1}^{m_2}\Sigma^m\right)$ takes a set~$S$ of~$n$ elements as input, and outputs representation~$M$ of length~$m_1 \leq m \leq m_2$.
%
The determinisitc query algorithm $\mathsc{Qry}\colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right)\times U \rightarrow \bits^*$ takes a representation $M$ and an element $x \in U$ and returns a bitstring.  
%
The deterministic \emph{update} algorithm $\mathsc{Up}\colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right)\times \bits^* \rightarrow \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right) \cup \{\bot\}$ takes a representation and an update string, and returns an updated representation or the distinguished symbol~$\bot$. 
%
The deterministic \emph{test} algorithm $\mathsc{Test} \colon \left(\bigcup_{m=m_1}^{m_2}\Sigma^m \right) \times U \rightarrow \bits$ takes a representation~$M$ and an element~$x \in U$ and returns a bit.
%
We assume that all $\mathsc{Rep},\mathsc{Qry},\mathsc{Up},\mathsc{Test}$ all have blackbox access to the functions $h_1,h_2,\ldots,h_k \in \mathcal{H}$, which we denote by writing~$\mathcal{H}$ as a superscript.

\tsnote{Not sure how to define correctness.  Intuitively, it should be something like, at any point in time, if $x\in S$ then $\Prob{\mathsc{Test}(M,x)=1} \geq 1-\epsilon$ for some correctness parameter~$\epsilon$, where the probablilty is over (at least) the coins of~$\mathsc{Hash}$. But perhaps this should be an adversarial game, in which case it is over the coins of the adversary, too...}

\heading{Discussion. } Our syntax captures settings in which there is some (potentially empty) initial set~$S$, whose representation~$M$ may undergo updates over time.  Effectively, this allows for expansion of~$S$ to a larger set, or even a multiset, as new elements ``arrive''.    The update algorithm is responsible for altering the current representation.  We allow it to take in a string that encodes inputs needed to carry out the updating.  For example, counting Bloom filters may receive update strings that encode $(x,c)$ where $x \in U$ is the element whose representation should be incremented, and $c \in \mathbb{N}$ is the amount of the increment.  Network applications, such as looking for heavy-hitters across TCP/IP streams seen by a router, may have $x = (\mathrm{IP_{src}},\mathrm{IP_{dst}})$, the source and destintation address of a packet, and~$c$ the number of bytes in the packet payload.  When necessary we will specify what is encoded in the update string, but will assume some implicit and fixed encoding scheme.

@@

An $(n,k,[m_1,m_2])$-filter with operations is a tuple  $B=(\mathsc{Hash},\mathsc{Rep},\mathsc{Qry}, \mathcal{F})$.  
The set $\mathcal{F}$ is the finite collection of allowable operations.  All operations are of the form 
$f: \mathbb{M}_{\mathcal{U}} \times \mathbb{M}_{\mathcal{U}} \rightarrow \mathbb{M}_{\mathcal{U}} \cup \{\bot\}$.  
%
The deterministic representation algorithm $\mathsc{Rep}\colon \mathbb{M}_\mathcal{U} \rightarrow \bigcup_{m=m_1}^{m_2}\Sigma^m$ takes a multiset~$S$, and outputs representation~$M$ of length~$m_1 \leq m \leq m_2$, or the distinguished symbol~$\bot$.  We assume that if the multiset~$S=\{(x_1,\ell_1),(x_2,\ell_2),\ldots,(x_t,\ell_t)\}$ is such that $n < \sum_{i=1}^t \ell_i$ then $\mathsc{Rep}(S)=\bot$.
%
The randomized hash-sampling algorithm~$\mathsc{Hash}$ is as before.
%
The determinisitic query algorithm $\mathsc{Qry}$... \tsnote{Not sure how to define this!  See my comment, below...}
